Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edited 4.3.1

When considering exponentiated quadratic and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$. With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
%
This is natural as with decreasing $\ell$, more higher frequency basis functions are needed. With increasing $c$, the smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
%
When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$ if $\ell$ or $c$ is changed.

@@

From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved. Following we show the numerical functions governing these relationships for every kernel.

@@

Notice that we establish a lower bound for $c$ set to 1.2.

>>>>

When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$. With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
%
This is natural as with decreasing $\ell$, more higher frequency basis functions are needed. With increasing $c$, the smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
%
When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$ if $\ell$ or $c$ is changed.

@@

When $c$ is large enough, there is near
linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ (see Section~\ref{subsec_theoretic_linearity}). To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the needed constant terms. We establish a lower bound for $c$ set to 1.2.

% From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved. Following we show the numerical functions governing these relationships for every kernel.

@@

These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$). Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions. If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.
